---
title: "Report of the results"
format: html
editor: visual
---

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

rm(list=ls())
library(readxl)
library(lavaan)
library(semTools)
library(psych)
library(ggplot2)
library(tidyr)
library(dplyr)
library(corrplot)
library(here)
library(reshape2)

library(gamlss)
library(gamlss.dist)  # LOGNO, qLOGNO, etc.
library(gamlss.add)   # smoothers like pb()

source(here("R/custom-functions.R"))
fi = c("rmsea","srmr","cfi","nnfi")

#df <- readxl::read_excel("data/Database_Time_16_05_25_PD_MI.xlsx")
df = data.frame(read_excel(here("data/Database_Time_16_05_25_PD_MI.xlsx")))
nrow(df)
df = df[!is.na(df$Class),] 
nrow(df)

dd = data.frame(read_excel(here("data/Data_Dictionary.xlsx")))
dd$Var_name = gsub("[ °-]",".",dd$Var_name)

ageBrack <- seq(from=6,to=12,by=0.25)

# perc_list <- seq(0.05,0.95,0.05) #
perc_list <- c(0.00135, 0.00383, 0.00982,
               0.02275, 0.04779, 0.09121,
               0.15866, 0.25249, 0.36944,
               .50,
               0.63056, 0.74751, 0.84134,
               0.90879, 0.95221, 0.97725,
               0.99018, 0.99617, 0.99865) # 19 punti

```

```{r}
#| label: scoring

df$OTm = apply(df[,paste0("OTm_",1:16)],1,sum)
df$QSTmt = apply(df[,paste0("QSTm_",1:10,"_t")],1,sum)
df$QSTmp = apply(df[,paste0("QSTm_",1:10,"_p")],1,sum)

dat <- df

```

```{r}
#| label: constructs
#| eval: false
#| echo: false

# list of constructs and ratios of complete observations
tb = as.data.frame(table(dd$ScaleName,dd$Construct))
tb = tb[tb$Freq!=0,1:2]
names(tb) = c("Scale","Construct")
tb$completeObs = NA
for(i in 1:nrow(tb)){
  x = dd$Var_name[dd$ScaleName==tb$Scale[i] & dd$Construct==tb$Construct[i]]; x = x[!is.na(x)]
  tb$completeObs[i] = round(mean(rowMeans(!is.na(df[,x]))),3)
}; tb
```

## Descriptives

```{r}
#| label: descriptives
length(unique(df$School)); length(unique(df$Class))
table(df$Grade)

table(df$Gender)
table(df$Gender)/700

table(df$ADHD_diagnosis)

table(df$DSA_diagnosis)
```

# Scales analysis

## Time sense parents

Parents responded to 10 items regarding their children sense of time. The items were on a 4-point Likert scale (0-3) and, as clear from @fig-qst-parents, the response distribution was extremely skewed with parents almost never reporting extremely high negative values.

```{r}
#| label: fig-qst-parents

var = "QSTpadua_parent"
items = dd$Var_name[dd$Level=="item"&dd$ScaleName==var]
itemsQstParent = items[!is.na(items)]

df_melted <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = itemsQstParent)

ggplot(df_melted, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable, ncol = 5)
```

### Factor structure

Given the properties of the Likert distribution, we thus fit a single-factor CFA using DWLS estimation for ordinal data. The model, however, does not have a great fit. Inspecting modification indices, it clearly appears that positive items (see also the distributions of items 1,2,4,5,6) are more similar than it should be.

```{r}
#| label: model-qst-parents
modelQstParent = paste0(gsub(" ","",var),"_lat=~",paste0(itemsQstParent,collapse="+"))
print(modelQstParent)
fitQstParent = cfa(model=modelQstParent, data=df, ordered=T)
fitQstParent@Data
fitMeasures(fitQstParent,fit.measures=fi)
modificationIndices(fitQstParent,sort.=T)[1:10,]
```

I added two covariances, but this does not help much. Modeling covariance across all the positive items gives a perfect fit, but loadings of positive items almost disappear.

```{r}
#| label: model-parents-mi

mi_line <- "
QSTp_4_p ~~ QSTp_5_p
QSTp_2_p ~~ QSTp_6_p
"

modelQstParentRev <- paste(c(modelQstParent, mi_line), collapse = "\n")
fitQstParentRev = cfa(model=modelQstParentRev, data=df, ordered=T, orthogonal = T)
fitMeasures(fitQstParentRev,fit.measures=fi)
# summary(fitQstParentRev,std=T)
```

An alternative solution might be fitting a two-factor model. This has optimal fit indices, however, the correlation between the two LV is very low (.14). Fitting a hierarchical model raises convergence issues.

```{r}
#| label: model-parents-two

model2 <- "
qstp1 =~ QSTp_1_p+QSTp_2_p+QSTp_4_p+QSTp_5_p+QSTp_6_p
qstp2 =~ QSTp_3_p+QSTp_7_p+QSTp_8_p+QSTp_9_p+QSTp_10_p
qstp1~~qstp2
# qst=~qstp1+qstp2
"

fitQstParent2 = cfa(model=model2, data=df, ordered=T, std.lv=T)
fitMeasures(fitQstParent2,fit.measures=fi)
# summary(fitQstParent2,std=T)
```

A final solution might be fitting a bifactor model. This has again optimal fit indices, however, loadings on the general factor are good for negative items but small \<.30 for positive items.

```{r}
#| label: model-parents-bi

mi_line2 <- "
qpos =~ QSTp_1_p+QSTp_2_p+QSTp_4_p+QSTp_5_p+QSTp_6_p
qnneg =~ QSTp_3_p+QSTp_7_p+QSTp_8_p+QSTp_9_p+QSTp_10_p
"
modelQstParentBi <- paste(c(modelQstParent, mi_line2), collapse = "\n")
fitQstParentBi = cfa(model=modelQstParentBi, data=df, ordered=T, orthogonal = T)
fitMeasures(fitQstParentBi,fit.measures=fi)

# summary(fitQstParentBi,std=T)
```

In any case, the questionnaire shows good reliability:

```{r}
#| label: parent-rel
#| warning: false
psych::alpha(df[,itemsQstParent], check.keys = TRUE)$total$raw_alpha
```

```{r}
bfi <- BifactorIndicesCalculator::bifactorIndices(
  fitQstParentBi)
bfi
```

### Scaling and age trend

The @fig-qst-parents-trend shows that parent-reported scores increase with age, but do not follow exactly a linear trend, probably given the scale limits.

```{r}
#| label: fig-qst-parents-trend
ggplot(df, aes(x = Age, y = QSTp_Total_parent)) +
  geom_smooth() +
  geom_point(alpha = .2) + 
  theme(text=element_text(size=18)) 

```

To better estimate standardized scores across age, we adopted gamlss and a Beta Binomial family to account for the ordinal nature of the data and the bounds.

```{r}
#| warning: false


dat_qstP = dat[!is.na(dat$QSTp_Total_parent),c("Code","Age","QSTp_Total_parent")]

fit_QSTp <- gamlss(
  cbind(QSTp_Total_parent, 30 - QSTp_Total_parent) ~ pbm(Age),
  sigma.fo = ~ pbm(Age),         
  family = BB,#BEINF,
  data   = dat_qstP
)
plot(fit_QSTp)

dat_qstP$mu <- predict(fit_QSTp, what = "mu", type = "response")
# Expected score (0..30)
dat_qstP$expected_score <- 30 * dat_qstP$mu
plot(dat_qstP$Age,dat_qstP$expected_score)

# z-score
dat_qstP$qstp_parent_z<-residuals(fit_QSTp, what = "z-scores")

norms_qstP <- matrix(NA_real_, 
                     nrow = length(perc_list), 
                     ncol = length(ageBrack))
colnames(norms_qstP) <- paste0("age-",ageBrack)
rownames(norms_qstP) <- perc_list

for (i in seq_along(ageBrack)) {
  pa <- predictAll(fit_QSTp,
                   newdata = data.frame(Age = ageBrack[i]),
                   type = "response")  # get mu, sigma on response scale
  mu_i    <- pa$mu
  sigma_i <- pa$sigma

  # Beta–binomial quantiles: returns integers in 0..bd
  norms_qstP[, i] <- qBB(p = perc_list, mu = mu_i, sigma = sigma_i, bd = 30)
}
```

## Time sense teachers

As for the parent questionnaire, teachers responded to 10 items on a 4-point Likert scale. In this case, however, we only have positively-keyed items and, with the exception of items 1, 6, and 10 (@fig-qst-teachers), most answers receive the highest score.

```{r}
#| label: fig-qst-teachers

# Questionnaire teachers ####
var = "QSTmilan_teacher"
items = dd$Var_name[dd$Level=="item"&dd$ScaleName==var]
itemsQstTeacher = items[!is.na(items)]

df_melted <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = itemsQstTeacher)

ggplot(df_melted, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable, ncol = 5)
```

### Factor structure

Differently from the parents' questionnaire, teachers' scale shows excellent fit to the data.

```{r}
#| label: model-qst-teachers
modelQstTeacher = paste0(gsub(" ","",var),"_lat=~",paste0(itemsQstTeacher,collapse="+"))
print(modelQstTeacher)
fitQstTeacher = cfa(model=modelQstTeacher, data=df, ordered=T)
fitQstTeacher@Data
fitMeasures(fitQstTeacher,fit.measures=fi)
```

The questionnaire also shows good reliability:

```{r}
#| label: teacher-rel
#| warning: false
psych::alpha(df[,itemsQstTeacher], check.keys = FALSE)$total$raw_alpha
```

### Scaling and age trend

The @fig-qst-teacher-trend shows that score increase with age, but do not follow exactly a linear trend, probably given the scale limits.

```{r}
#| label: fig-qst-teacher-trend
ggplot(df, aes(x = Age, y = QSTp_Total_teacher)) +
  geom_smooth() +
  geom_point(alpha = .2) + 
  theme(text=element_text(size=18)) 

```

We thus follow the same gamlss approach as before.

```{r}
dat <- df

dat_qstT = dat[!is.na(dat$QSTp_Total_teacher),c("Code","Age","QSTp_Total_teacher")]

fit_QSTt <- gamlss(
  cbind(QSTp_Total_teacher, 30 - QSTp_Total_teacher) ~ pbm(Age),
  sigma.fo = ~ pbm(Age),         
  family = BB,#BEINF,
  data   = dat_qstT
)
plot(fit_QSTt)

dat_qstT$mu <- predict(fit_QSTt, what = "mu", type = "response")
# Expected score (0..30)
dat_qstT$expected_score <- 30 * dat_qstT$mu
plot(dat_qstT$Age,dat_qstT$expected_score)

# z-score
dat_qstT$qstp_teacher_z<-residuals(fit_QSTt, what = "z-scores")

norms_qstT <- matrix(NA_real_, 
                     nrow = length(perc_list), 
                     ncol = length(ageBrack))
colnames(norms_qstT) <- paste0("age-",ageBrack)
rownames(norms_qstT) <- perc_list

for (i in seq_along(ageBrack)) {
  pa <- predictAll(fit_QSTt,
                   newdata = data.frame(Age = ageBrack[i]),
                   type = "response")  # get mu, sigma on response scale
  mu_i    <- pa$mu
  sigma_i <- pa$sigma

  # Beta–binomial quantiles: returns integers in 0..bd
  norms_qstT[, i] <- qBB(p = perc_list, mu = mu_i, sigma = sigma_i, bd = 30)
}

```

## Time knowledge and orientation

The time orientation questionnaire is composed of 16 questions about time knowledge and orientation. The first 10 items measure time knowledge, while the last 6 measure time orientation. Items are scored as 1 if correct or 0 if wrong (@fig-time-orientation).

```{r}
#| label: fig-time-orientation

# Time Orientation #### 
var = "OTm"
items = dd$Var_name[dd$Level=="item"&dd$ScaleName==var]
itemsTimeOt = items[!is.na(items)]

df_melted <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = itemsTimeOt)

ggplot(df_melted, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable, ncol = 8)

```

### Factor structure

The overall model shows good, but not perfect fit to the data. Looking at the modification indices, items 12 and 14 might be problematic.

```{r}
#| label: model-time-orientation-tot
modelTimeOt = paste0(gsub(" ","",var),"_lat=~",paste0(itemsTimeOt,collapse="+"))
print(modelTimeOt)
fitTimeOt = cfa(model=modelTimeOt, data=df, ordered=T)
fitTimeOt@Data
fitMeasures(fitTimeOt,fit.measures=fi)
modificationIndices(fitTimeOt,sort.=T)[1:10,] # Residui correlati item 12 e 14 - unici con risposta sì-no
```

Adding the covariance between items 12 and 14 improves model fit:

```{r}
#| label: model-time-orientation-tot-mi
mi_line <- "
OTm_12	~~	OTm_14
"

modelTimeOtRev <- paste(c(modelTimeOt, mi_line), collapse = "\n")
fitTimeOtRev = cfa(model=modelTimeOtRev, data=df, ordered=T)
fitMeasures(fitTimeOtRev,fit.measures=fi)
```

The scale has also decent internal consistency for a 0-1 scale:

```{r}
#| label: time-orientation-tot-rel
#| warning: false
psych::alpha(df[,itemsTimeOt], check.keys = TRUE)$total$raw_alpha
```

And a good one considering ordinal reliability:

```{r}
#| warning: false
#| message: false
semTools::reliability(fitTimeOtRev)[2,]
```

### Scaling and age trend

The @fig-qst-teacher-trend shows that score increase with age following an approximately linear trend before 10 and a smaller increase later on.

```{r}
#| label: fig-time-orientation-trend
ggplot(df, aes(x = Age, y = OTm)) +
  geom_smooth() +
  geom_point(alpha = .2) + 
  theme(text=element_text(size=18)) 

```

Being these items again on a Likert scale (0-1), we adopt again a BB family in gamlss.

```{r}
dat_otm = dat[!is.na(dat$OTm),c("Code","Age","OTm")]

fit_OTm <- gamlss(
  cbind(OTm,16-OTm) ~ pbm(Age),
  sigma.fo = ~ pbm(Age),         
  family = BB,
  data   = dat_otm
)
plot(fit_OTm)

dat_otm$mu    <- fitted(fit_OTm, what = "mu", type = "response") 
# Expected score (0..30)
dat_otm$expected_score <- 16 * dat_otm$mu
plot(dat_otm$Age,dat_otm$expected_score)

# z-score
dat_otm$OTm_z<-residuals(fit_OTm, what = "z-scores")

norms_OTm  <- matrix(NA_real_, 
                     nrow = length(perc_list), 
                     ncol = length(ageBrack))
colnames(norms_OTm) <- paste0("age-",ageBrack)
rownames(norms_OTm) <- perc_list

for (i in seq_along(ageBrack)) {
  pa <- predictAll(fit_OTm,
                   newdata = data.frame(Age = ageBrack[i]),
                   type = "response")  # get mu, sigma on response scale
  mu_i    <- pa$mu
  sigma_i <- pa$sigma

  # Beta–binomial quantiles: returns integers in 0..bd
  norms_OTm[, i] <- qBB(p = perc_list, mu = mu_i, sigma = sigma_i, bd = 16)
}

```

#### Time knowledge

As before, we fit again a CFA on this data using DWLS estimation for ordered categories (0/1). The model fits the data very well.

```{r}
#| label: model-time-know
items_k = itemsTimeOt[1:10]

modelTimeKnow = paste0(gsub(" ","",var),"_latk=~",paste0(items_k,collapse="+"))
print(modelTimeKnow)
fitTimeKnow = cfa(model=modelTimeKnow, data=df, ordered=T)
fitTimeKnow@Data
fitMeasures(fitTimeKnow,fit.measures=fi)
```

#### Time orientation

The time orientation scale has perfect fit to the data (covariance between item 12 and 14 is included).

```{r}
#| label: model-time-orientation
items_o = itemsTimeOt[11:16]
modelTimeOt = paste0(gsub(" ","",var),"_lato=~",paste0(items_o,collapse="+"))
mi_line <- "
OTm_12	~~	OTm_14
"

modelTimeOtRev <- paste(c(modelTimeOt, mi_line), collapse = "\n")
print(modelTimeOtRev)
fitTimeOtRev = cfa(model=modelTimeOtRev, data=df, ordered=T)
fitTimeOtRev@Data
fitMeasures(fitTimeOtRev,fit.measures=fi)
```

##### Two-factor model

A two factor model shows good fit to the data, but the two scores are highly correlated. We thus keep one factor only.

```{r}
#| label: model-time-2factor
model2 <- paste(c(modelTimeOt,modelTimeKnow), collapse = "\n")
fitTime2 = cfa(model=model2, data=df, ordered=T)
fitmeasures(fitTime2,fit.measures = fi)

standardizedSolution(fitTime2)[51,"est.std"]
```

## Time discrimination and estimation

The two tests measuring time discrimination and estimation cannot be modeled using factor analyses as they are either composed by two items (time estimation) or return a single score based on the lowest discrimination ratio calculated with adaptive processes.

### Time discrimination

For time discrimination, the distribution of the scores is again very skewed (see @fig-td), with most children scoring between 1 and 1.5.

```{r}
#| label: fig-td
# dd$Var_name[dd$Construct == "Time Discrimination"]

ggplot(df, aes(x = RatioTD)) +
  geom_histogram(binwidth = .02, 
                 fill = "orange", color = "black") 
```

#### Scaling and age trend

The @fig-td-trend shows that score increase with age (negative scores), but do not follow exactly a linear trend.

```{r}
#| label: fig-td-trend
ggplot(df, aes(x = Age, y = RatioTD)) +
  geom_smooth() +
  geom_point(alpha = .2) + 
  theme(text=element_text(size=18)) +
  coord_cartesian(ylim = c(1,3))

```

To accomodate the distribution of this data, given by a ratio, we firstly transform it to a proportion and then use a Box–Cox t (original scale) transformatio. This is based on a Box–Cox transformation of the response variable, combined with a Student’s t distribution for the transformed values.

```{r}
dat_td = dat[!is.na(dat$RatioTD),c("Code","Age","RatioTD")]

dat_td$TD_prop <- (dat_td$RatioTD - 1) / (3.005 - 1)   # now in [0,1]

# fit_TD <- gamlss(
#   TD_prop ~ pbm(Age),
#   sigma.fo = ~ pbm(Age),
#   family = BE,     # Beta
#   data = dat_td
# )
# plot(fit_TD)
# 
# 
# dat_td$TD_prop <- (dat_td$RatioTD - 1) / (3.004 - 1)   # now in [0,1]
# 
# fit_TD <- gamlss(
#   TD_prop ~ cs(Age),
#   sigma.fo = ~ cs(Age),
#   nu.fo   = ~ 1,#pbm(Age),   # probability of 0 (maps to Ratio=1)
#   tau.fo  = ~ cs(Age),   # probability of 1 (maps to Ratio=3)
#   family  = BEINF,       # Beta inflated at 0/1 because many have 1 or 3
#   data    = dat_td
# )
# plot(fit_TD)

# It is based on a Box–Cox transformation of the response variable, combined with a Student’s t distribution for the transformed values.

fit_LMS <- gamlss(
  TD_prop ~ pb(Age),
  sigma.fo = ~ pb(Age),
  nu.fo    = ~ 1,#pb(Age),   # skewness
  tau.fo   = ~ 1,#pb(Age),   # kurtosis
  family   = BCTo,
  data     = dat_td
)
plot(fit_LMS)

dat_td$mu    <- fitted(fit_LMS, what = "mu", type = "response") 

# Expected score (0..30)
dat_td$expected_score <- 1 + dat_td$mu * (3.004-1)
plot(dat_td$Age,dat_td$expected_score)

# z-score
dat_td$TD_z<-residuals(fit_LMS, what = "z-scores")

norms_TD  <- matrix(NA_real_, 
                    nrow = length(perc_list), 
                    ncol = length(ageBrack))
colnames(norms_TD) <- paste0("age-",ageBrack)
rownames(norms_TD) <- perc_list

for (i in seq_along(ageBrack)) {
  pa <- predictAll(fit_LMS,
                   newdata = data.frame(Age = ageBrack[i]),
                   type = "response")  # get mu, sigma on response scale
  mu_i    <- pa$mu
  sigma_i <- pa$sigma
  nu_i    <- pa$nu
  tau_i   <- pa$tau

  norms_TD[, i] <- qBCTo(p = perc_list, 
                         mu = mu_i, sigma = sigma_i, nu = nu_i, tau = tau_i)
}
norms_TD <- 1 + norms_TD * (3.004-1)
```

### Time estimation

For time estimation, the score is given by the absolute difference (in seconds) from the correct answer. The distribution of the scores is again very skewed (see @fig-te).

```{r}
#| label: fig-te
# Time Estimation ####
# dd$Var_name[dd$Construct == "Time Estimation"]
teItems <- c("TE_Barca_dd","TE_Ladro_dd","TEdd_Mean")

df_melted <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = teItems)

ggplot(df_melted, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable)
```

The correlations between the two items is very high, suggesting that common factors explained children performance in both items.

```{r}
#| label: te-cor
cor(df$TE_Barca_dd,df$TE_Ladro_dd, use = "complete.obs")
```

#### Scaling and age trend

The @fig-te-trend shows that score increase with age (lower discrepancies).

```{r}
#| label: fig-te-trend
dfte <- melt(subset(df, select = c(Age,TE_Barca_dd,TE_Ladro_dd)),
             id.vars = "Age")

ggplot(dfte, aes(x = Age, y = value)) +
  geom_smooth() +
  geom_point(alpha = .2) + 
  theme(text=element_text(size=18)) +
  coord_cartesian(ylim = c(0,50)) +
  facet_wrap(~variable)


```

To accomodate the distribution of the data, we use a zero-adjusted gamma distribution and standardize both items separately.

##### Barca

```{r}
#| label: scaling-barca
dat_teB <- subset(dat, !is.na(TE_Barca_dd), select = c(Code, Age, TE_Barca_dd))

fit_teB <- gamlss(
  TE_Barca_dd ~ pbm(-Age),      # monotone decreasing mean
  sigma.fo   = ~ pbm(Age),       # allow heteroscedasticity
  family     = ZAGA,           # zero-adjusted gamma
  data       = dat_teB
)

dat_teB$muB    <- fitted(fit_teB, what = "mu", type = "response") 
plot(dat_teB$Age,dat_teB$muB)

# z-score
dat_teB$teB_z<-residuals(fit_teB, what = "z-scores")

norms_teB  <- matrix(NA_real_, 
                     nrow = length(perc_list), 
                     ncol = length(ageBrack))
colnames(norms_teB) <- paste0("age-",ageBrack)
rownames(norms_teB) <- perc_list

for (i in 1:length(ageBrack)) {
  pA <- predict(fit_teB, what="mu",    type="response", 
                 newdata = data.frame(Age = ageBrack[i]))
  norms_teB[,i] <- qZAGA(p = perc_list, mu = pA)
}

```

##### Ladro

```{r}
#| label: scaling-ladro
dat_teL <- subset(dat, !is.na(TE_Ladro_dd), select = c(Code, Age, TE_Ladro_dd))

fit_teL <- gamlss(
  TE_Ladro_dd ~ pbm(-Age),      # monotone decreasing mean
  sigma.fo   = ~ pbm(Age),       # allow heteroscedasticity
  family     = ZAGA,           # zero-adjusted gamma
  data       = dat_teL
)

dat_teL$muL    <- fitted(fit_teL, what = "mu", type = "response") 
plot(dat_teL$Age,dat_teL$muL)

# z-score
dat_teL$teL_z<-residuals(fit_teL, what = "z-scores")

norms_teL  <- matrix(NA_real_, 
                     nrow = length(perc_list), 
                     ncol = length(ageBrack))
colnames(norms_teL) <- paste0("age-",ageBrack)
rownames(norms_teL) <- perc_list

for (i in 1:length(ageBrack)) {
  pA <- predict(fit_teL, what="mu",    type="response", 
                 newdata = data.frame(Age = ageBrack[i]))
  norms_teL[,i] <- qZAGA(p = perc_list, mu = pA)
}

```

To test the convergence of the two items, we transformed the scores in z scores standardized by age and calculated their correlation.

```{r}
#| label: te-cors

d_te <- merge(subset(dat_teB, select = c(Code,Age, teB_z, muB)),
              subset(dat_teL, select = c(Code,teL_z, muL))
              )
d_te <- d_te %>%
  mutate(across(where(is.numeric), ~ replace(., is.infinite(.), NA_real_)))

cor(d_te$teB_z,d_te$teL_z, use = "pairwise.complete")

# cor(log(df$TE_Barca) - log(30), log(df$TE_Ladro - log(30)), use = "pairwise.complete")

```

We then computed the mean score for each participant.

```{r}
d_te$total_z <- rowMeans(subset(d_te, select = c(teB_z,teL_z)))
```

## Time reproduction

For time reproduction, children perceived 12 time sequences and reproduced them. The scores are calculated as the absolute value of the ratio between reproduced and correct time.

This, however, generates extremely skewed distributions of the scores (see @fig-tr-raw). Hence, we transformed the items in z scores before analysis.

```{r}
#| label: fig-tr-raw
#| cache: true
# dd$Var_name[dd$Construct == "Time Reproduction"]
items = dd$Var_name[dd$ScaleName=="TR"&dd$Level=="item"]
items = items[!is.na(items) & grepl("PercDevAbs_TR_",items)]

df_melted <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = items)

ggplot(df_melted, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable, ncol = 4)
```

#### Scaling and age trend

The @fig-tr-trend shows that scores decrease almost linearly with age. We thus proceeded to normalize them all for age using gamlss and a BCTo family again.

```{r}
#| label: fig-tr-trend
ggplot(df_melted, aes(x = Age, y = value)) +
  geom_smooth() +
  geom_point(alpha = .2) + 
  theme(text=element_text(size=18)) +
  coord_cartesian(ylim = c(0,200)) +
  facet_wrap(~variable, ncol = 4)
```

```{r}
dat_tr <- dat[, c("Code","Age", items)]
dat_tr <- dat_tr[complete.cases(dat_tr),]

fit_tr2 <- gamlss(
  PercDevAbs_TR_2 ~ pb(Age),
  sigma.fo = ~ pb(Age),
  family = BCTo,
  data = dat_tr
)


pa <- predictAll(fit_tr2,
                   newdata = data.frame(Age = ageBrack[i]),
                   type = "response")  # get mu, sigma on response scale
    mu_i    <- pa$mu
    sigma_i <- pa$sigma
    nu_i    <- pa$nu
    tau_i   <- pa$tau
    
    norms_TD[, i] <- qBCTo(p = perc_list,
                           mu = mu_i, sigma = sigma_i, nu = nu_i, tau = tau_i)

fitList <- list()

for (var in items) {
  
  # Build formula dynamically
  fmla <- as.formula(paste(var, "~ pb(Age)"))
  
  # Fit the model
  fitList[[var]] <- gamlss(
    formula = fmla,
    sigma.fo = ~ pb(Age),
    family   = BCTo,      # or IG / GA / BCCG, depending on what you decided
    data     = dat_tr,
    trace    = FALSE
  )
  
  # Compute z-scores
  z_name <- paste0(sub("PercDevAbs_", "", var), "_z")   # e.g., "TR_2_z"
  dat_tr[[z_name]] <- residuals(fitList[[var]], what = "z-scores")
  
}

TR_norms_list <- list()

for (k in 1:length(items)) {
  TR_norms_list[[k]] <- matrix(NA_real_, 
                     nrow = length(perc_list), 
                     ncol = length(ageBrack))
  colnames(TR_norms_list[[k]]) <- paste0("age-",ageBrack)
  rownames(TR_norms_list[[k]]) <- perc_list
  
  fitTR <- fitList[[k]]
  
  for (i in 1:length(ageBrack)) {
  
    pa <- predictAll(fitTR,
                     newdata = data.frame(Age = ageBrack[i]),
                     type = "response")  # get mu, sigma on response scale
    mu_i    <- pa$mu
    sigma_i <- pa$sigma
    nu_i    <- pa$nu
    tau_i   <- pa$tau
    
    TR_norms_list[[k]][, i] <- qBCTo(p = perc_list,
                                     mu = mu_i, sigma = sigma_i, 
                                     nu = nu_i, tau = tau_i)
  }
}

```

The resulting scores are well correlated with each other (see @fig-tr-cors)

```{r}
#| label: fig-tr-cors
dat_tr$tr_z <- rowSums(subset(dat_tr,select = TR_2_z:TR_12_z))

corrplot(round(cor(subset(dat_tr,select = TR_2_z:TR_12_z)),2),
         method = "color",  
         type = "full",    
         addCoef.col = "black",
         tl.col = "black")

```

We thus proceed with factor analysis to confirm the factorial validity of the scale.

### Factor structure

The CFA model fit using the normalized items shows, as expected, a good fit to the data.

```{r}
#| label: model-tr
model = paste("TR_lat =~",paste(paste0("TR_",2:12,"_z"),collapse="+"))
model
fit = cfa(model=model, data=dat_tr)
fit@Data
fitMeasures(fit,fit.measures=fi)
```

The items are also reliable:

```{r}
#| label: rel-tr
#| warning: false

psych::alpha(dat_tr[,paste0("TR_",2:12,"_z")])$total$raw_alpha

```

# Convergence of the scales

Finally, we estimated whether the scales converge to ensure that all tap into similar constructs.

## Raw correlations

With the exception of the time estimation score, the time scales are all correlated between each other and with participants' age, but not with gender (@fig-raw-cor). The common age effect highlights the need for age norming to remove age-related covariance.

```{r}
#| label: fig-raw-cor
dcor <- subset(df, select = c(QSTp_Total_parent, QSTm_Total_teacher,
                              OTm, RatioTD,
                              TEdd_Mean, AverageDevAbs_TR,
                              # Grade, 
                              Gender,Age))
# dcor$AverageDevAbs_TR <- normalize(dcor$AverageDevAbs_TR)
corrplot::corrplot(cor(dcor, use = "pairwise.complete"), addCoef.col = "black")
```

## z-score correlations

We thus compute correlations between the z scores obtained through the process (@fig-cors-z).

```{r}
#| label: fig-cors-z
library(dplyr)

d_z <- dat_qstP %>% select(Code, qstp_parent_z) %>%
  full_join(dat_qstT %>% select(Code, qstp_teacher_z), by = "Code") %>%
  full_join(dat_otm  %>% select(Code, OTm_z),          by = "Code") %>%
  full_join(dat_td   %>% select(Code, TD_z),           by = "Code") %>%
  full_join(d_te %>% select(Code,total_z),  by = "Code") %>% #teB_z,teL_z,
  full_join(dat_tr %>% select(Code, tr_z), by = "Code")
names(d_z)[6] <- "TE_z"
names(d_z)[7] <- "TR_z"
corrplot::corrplot(cor(d_z[,-1], use = "pairwise.complete"), addCoef.col = "black")

```

## Factor structure

We may view the different factors measured as indicator of a general factor of time. This might be modeled as a CFA.

```{r}
#| label: general-factor-model
g_model = paste("g =~",paste(names(d_z)[-1],collapse="+"))
g_model
fit = cfa(model=g_model, data=d_z, missing = "fiml",std.lv=T)
fit@Data
fitMeasures(fit,fit.measures=fi)
modificationIndices(fit,sort.=T)[1:10,]
```

The model shows acceptable, yet quite low, fit indices. MI suggest that time discrimination and time reproduction scores may be more similar between each other than with other factors. This makes sense as the two have quite similar requests. With this specification, the model shows excellent fit to the data.

```{r}
#| label: general-factor-model-mi
mi_line <- "
TD_z	~~	TR_z
"

g_model <- paste(c(g_model, mi_line), collapse = "\n")
fit = cfa(model=g_model, data=d_z, missing = "fiml",std.lv=T)
fitMeasures(fit,fit.measures=fi)
```

```{r}
#| label: general-factor-model-mi-summary

summary(fit,std=T)
```

```{r, eval=TRUE}
tabList <- list(norms_OTm,norms_qstP,norms_qstT,
                norms_TD,norms_teB,norms_teL,
                TR_norms_list)
names(tabList) <- c("OTm", "QSTp", "QSTt", "TD", "teB", "teL","TR")
saveRDS(tabList,here::here("scoring/normPercTabs.RDS"))
```
