---
title: "report"
format: html
editor: visual
---

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

rm(list=ls())
library(readxl)
library(lavaan)
library(semTools)
library(psych)
library(ggplot2)
library(tidyr)
library(dplyr)
library(corrplot)
library(here)
library(reshape2)
source(here("R/custom-functions.R"))
fi = c("rmsea","srmr","cfi","nnfi")

df = data.frame(read_excel(here("data/Database_Time_16.05.25_PD&MI.xlsx"))); nrow(df)
df = df[!is.na(df$Class),]; nrow(df)

dd = data.frame(read_excel(here("data/Data_Dictionary.xlsx")))
dd$Var_name = gsub("[ °-]",".",dd$Var_name)

```

```{r}
#| label: scoring

df$OTm = apply(df[,paste0("OTm_",1:16)],1,sum)
df$QSTmt = apply(df[,paste0("QSTm_",1:10,"_t")],1,sum)
df$QSTmp = apply(df[,paste0("QSTm_",1:10,"_p")],1,sum)

```

```{r}
#| label: constructs
#| eval: false
#| echo: false

# list of constructs and ratios of complete observations
tb = as.data.frame(table(dd$ScaleName,dd$Construct))
tb = tb[tb$Freq!=0,1:2]
names(tb) = c("Scale","Construct")
tb$completeObs = NA
for(i in 1:nrow(tb)){
  x = dd$Var_name[dd$ScaleName==tb$Scale[i] & dd$Construct==tb$Construct[i]]; x = x[!is.na(x)]
  tb$completeObs[i] = round(mean(rowMeans(!is.na(df[,x]))),3)
}; tb
```

## Descriptives

```{r}
#| label: descriptives
length(unique(df$School)); length(unique(df$Class))
table(df$Grade)
```

# Scales analysis

## Time sense parents

Parents responded to 10 items regarding their children sense of time. The items were on a 4-point Likert scale (0-3) and, as clear from @fig-qst-parents, the response distribution was extremely skewed with parents almost never reporting extremely high negative values.

```{r}
#| label: fig-qst-parents

var = "QSTpadua_parent"
items = dd$Var_name[dd$Level=="item"&dd$ScaleName==var]
itemsQstParent = items[!is.na(items)]

df_melted <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = itemsQstParent)

ggplot(df_melted, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable, ncol = 5)
```

Given the properties of the Likert distribution, we thus fit a single-factor CFA using DWLS estimation for ordinal data.

```{r}
#| label: model-qst-parents
modelQstParent = paste0(gsub(" ","",var),"_lat=~",paste0(itemsQstParent,collapse="+"))
print(modelQstParent)
fitQstParent = cfa(model=modelQstParent, data=df, ordered=T)
fitQstParent@Data
fitMeasures(fitQstParent,fit.measures=fi)
modificationIndices(fitQstParent,sort.=T)[1:10,]
```

The model, however, does not have a great fit. Inspecting modification indices, it clearly appears that positive items (see also the distributions of items 1,2,4,5,6) are more similar than it should be.

I added two covariances, but this does not help much. Modeling covariance across all the positive items gives a perfect fit, but loadings of positive items almost disappear.

```{r}
#| label: model-parents-mi

mi_line <- "
QSTp_4_p ~~ QSTp_5_p
QSTp_2_p ~~ QSTp_6_p
"
# mi_line <- "
# QSTp_1_p ~~ QSTp_2_p+QSTp_4_p+QSTp_5_p+QSTp_6_p
# QSTp_2_p ~~ QSTp_4_p+QSTp_5_p+QSTp_6_p
# QSTp_4_p ~~ QSTp_5_p+QSTp_6_p
# QSTp_5_p ~~ QSTp_6_p
# 
# QSTp_3_p ~~ QSTp_7_p+QSTp_8_p+QSTp_9_p+QSTp_10_p
# QSTp_7_p ~~ QSTp_8_p+QSTp_9_p+QSTp_10_p
# QSTp_8_p ~~ QSTp_9_p+QSTp_10_p
# QSTp_9_p ~~ QSTp_10_p
# "

modelQstParentRev <- paste(c(modelQstParent, mi_line), collapse = "\n")
fitQstParentRev = cfa(model=modelQstParentRev, data=df, ordered=T, orthogonal = T)
fitMeasures(fitQstParentRev,fit.measures=fi)
# summary(fitQstParentRev,std=T)
```

An alternative solution might be fitting a two-factor model. This has optimal fit indices, however, the correlation between the two LV is very low (.14). Fitting a hierarchical model raises convergence issues.

```{r}
#| label: model-parents-two

model2 <- "
qstp1 =~ QSTp_1_p+QSTp_2_p+QSTp_4_p+QSTp_5_p+QSTp_6_p
qstp2 =~ QSTp_3_p+QSTp_7_p+QSTp_8_p+QSTp_9_p+QSTp_10_p
qstp1~~qstp2
# qst=~qstp1+qstp2
"

fitQstParent2 = cfa(model=model2, data=df, ordered=T, std.lv=T)
fitMeasures(fitQstParent2,fit.measures=fi)
# summary(fitQstParent2,std=T)
```

In any case, the questionnaire shows good reliability:

```{r}
#| label: parent-rel
#| warning: false
psych::alpha(df[,itemsQstParent], check.keys = TRUE)$total$raw_alpha
```

## Time sense teachers

As for the parent questionnaire, teachers responded to 10 items on a 4-point Likert scale. In this case, however, we only have positively-keyed items and, with the exception of items 1, 6, and 10 (@fig-qst-teachers), most answers receive the highest score.

```{r}
#| label: fig-qst-teachers

# Questionnaire teachers ####
var = "QSTmilan_teacher"
items = dd$Var_name[dd$Level=="item"&dd$ScaleName==var]
itemsQstTeacher = items[!is.na(items)]

df_melted <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = itemsQstTeacher)

ggplot(df_melted, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable, ncol = 5)
```

Differently from the parents' questionnaire, teachers' scale shows excellent fit to the data.

```{r}
#| label: model-qst-teachers
modelQstTeacher = paste0(gsub(" ","",var),"_lat=~",paste0(itemsQstTeacher,collapse="+"))
print(modelQstTeacher)
fitQstTeacher = cfa(model=modelQstTeacher, data=df, ordered=T)
fitQstTeacher@Data
fitMeasures(fitQstTeacher,fit.measures=fi)
```

The questionnaire also shows good reliability:

```{r}
#| label: teacher-rel
#| warning: false
psych::alpha(df[,itemsQstTeacher], check.keys = FALSE)$total$raw_alpha
```

## Time reproduction

For time reproduction, children perceived 12 time sequences and reproduced them. The scores are calculated as the absolute value of the ratio between reproduced and correct time.

This, however, generate an extremely skewed distribution of the scores (see @fig-tr-raw). Hence, we transformed the items before analysis.

```{r}
#| label: fig-tr-raw
#| cache: true
# dd$Var_name[dd$Construct == "Time Reproduction"]
items = dd$Var_name[dd$ScaleName=="TR"&dd$Level=="item"]
items = items[!is.na(items) & grepl("PercDevAbs_TR_",items)]

df_melted <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = items)

ggplot(df_melted, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable, ncol = 4)
```

Specifically, a quantile transformation was used to normalize the data (see @fig-tr-norm).

```{r}
#| label: fig-tr-norm

itemsNorm = paste0("TR",2:12)
df[,itemsNorm] = NA
for(i in 1:length(items)) df[,itemsNorm[i]] = normalize(abs(df[,items[i]]))

df_melted <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = itemsNorm)

ggplot(df_melted, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable, ncol = 4)
```

The CFA model fit using the normalized items shows, as expected, a good fit to the data.

```{r}
#| label: model-tr
model = paste("TR_lat =~",paste(itemsNorm,collapse="+"))
model
fit = cfa(model=model, data=df)
fit@Data
fitMeasures(fit,fit.measures=fi)
# modificationIndices(fit,sort.=T)[1:10,]
```

The items are also reliable:

```{r}
#| label: rel-tr
#| warning: false

psych::alpha(df[,itemsNorm])$total$raw_alpha

```

And well correlated (both when normalized \[above the diagonal\] and raw \[below\], see @fig-tr-cor):

```{r}
#| label: fig-tr-cor
corTabTrNorm <- cor(df[,itemsNorm],use="pairwise.complete",method="pearson")
corTabTrRaw <- cor(df[,items],use="pairwise.complete",method="spearman")
corTabTr <- corTabTrRaw  # start with raw (lower triangle)
corTabTr[upper.tri(corTabTr)] <- corTabTrNorm[upper.tri(corTabTrNorm)]
corrplot(corTabTr,
         method = "color",  
         type = "full",     
         addCoef.col = "black", 
         tl.col = "black")
```

## Time knowledge and orientation

The time orientation questionnaire is composed of 16 questions about time knowledge and orientation. The first 10 items measure time knowledge, while the last 6 measure time orientation. Items are scored as 1 if correct or 0 if wrong (@fig-time-orientation).

```{r}
#| label: fig-time-orientation

# Time Orientation #### 
var = "OTm"
items = dd$Var_name[dd$Level=="item"&dd$ScaleName==var]
itemsTimeOt = items[!is.na(items)]

df_melted <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = itemsTimeOt)

ggplot(df_melted, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable, ncol = 8)

```

```{r}
#| label: model-time-orientation-tot
modelTimeOt = paste0(gsub(" ","",var),"_lat=~",paste0(itemsTimeOt,collapse="+"))
print(modelTimeOt)
fitTimeOt = cfa(model=modelTimeOt, data=df, ordered=T)
fitTimeOt@Data
fitMeasures(fitTimeOt,fit.measures=fi)
modificationIndices(fitTimeOt,sort.=T)[1:10,] # Residui correlati item 12 e 14 - unici con risposta sì-no
```

Adding the covariance between items 12 and 14 improves model fit:

```{r}
#| label: model-time-orientation-tot-mi
mi_line <- "
OTm_12	~~	OTm_14
"

modelTimeOtRev <- paste(c(modelTimeOt, mi_line), collapse = "\n")
fitTimeOtRev = cfa(model=modelTimeOtRev, data=df, ordered=T)
fitMeasures(fitTimeOtRev,fit.measures=fi)
```

The scale has also decent internal consistency for a 0-1 scale:

```{r}
#| label: time-orientation-tot-rel
#| warning: false
psych::alpha(df[,itemsTimeOt], check.keys = TRUE)$total$raw_alpha
```

And a good one considering ordinal reliability:

```{r}
#| warning: false
#| message: false
semTools::reliability(fitTimeOtRev)[2,]
```


## Time knowledge

As before, we fit again a CFA on this data using DWLS estimation for ordered categories (0/1). The model fits the data very well.

```{r}
#| label: model-time-know
items_k = itemsTimeOt[1:10]

modelTimeKnow = paste0(gsub(" ","",var),"_lat=~",paste0(items_k,collapse="+"))
print(modelTimeKnow)
fitTimeKnow = cfa(model=modelTimeKnow, data=df, ordered=T)
fitTimeKnow@Data
fitMeasures(fitTimeKnow,fit.measures=fi)
```

## Time orientation

The time orientation scale has perfect fit to the data (covariance between item 12 and 14 is included).

```{r}
#| label: model-time-orientation
items_o = itemsTimeOt[11:16]
modelTimeOt = paste0(gsub(" ","",var),"_lat=~",paste0(items_o,collapse="+"))
mi_line <- "
OTm_12	~~	OTm_14
"

modelTimeOtRev <- paste(c(modelTimeOt, mi_line), collapse = "\n")
print(modelTimeOtRev)
fitTimeOtRev = cfa(model=modelTimeOtRev, data=df, ordered=T)
fitTimeOtRev@Data
fitMeasures(fitTimeOtRev,fit.measures=fi)
```

## Time discrimination and estimation

The two tests measuring time discrimination and estimation cannot be modeled using factor analyses as they are either composed by two items (time estimation) or return a single score based on the lowest discrimination ratio calculated with adaptive processes.

### Time discrimination

For time discrimination, the distribution of the scores is again very skewed (see @fig-td), with most children scoring between 1 and 1.5.

```{r}
#| label: fig-td
# dd$Var_name[dd$Construct == "Time Discrimination"]

ggplot(df, aes(x = RatioTD)) +
  geom_histogram(binwidth = .02, 
                 fill = "orange", color = "black") 
```

### Time estimation

For time estimation, the score is given by the absolute difference (in seconds) from the correct answer. The distribution of the scores is again very skewed (see @fig-te).

```{r}
#| label: fig-te
# Time Estimation ####
# dd$Var_name[dd$Construct == "Time Estimation"]
teItems <- c("TE_Barca_dd","TE_Ladro","TEdd_Mean")

df_melted <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = teItems)

ggplot(df_melted, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable)
```

The correlations between the two items is very high, suggesting that common factors explained children performance in both items.

```{r}
#| label: te-cor
cor(df$TE_Barca_dd,df$TE_Ladro_dd, use = "complete.obs")
```

# Convergence of the scales

## Correlations

With the exception of the time estimation score, the time scales are all correlated between each other and with participants' grade, but not with gender (@fig-raw-cor).

```{r}
#| label: fig-raw-cor
dcor <- subset(df, select = c(QSTp_Total_parent, QSTm_Total_teacher,
                              OTm, RatioTD,
                              TEdd_Mean, AverageDevAbs_TR,
                              Grade, Gender,Age))
# dcor$AverageDevAbs_TR <- normalize(dcor$AverageDevAbs_TR)
corrplot::corrplot(cor(dcor, use = "pairwise.complete"), addCoef.col = "black")
```

However, the scales are generally skewed (@fig-raw-tot). Hence, we normalized them (see @fig-norm-tot) and calculated the correlations again.

```{r}
#| label: fig-raw-tot
df_melted_raw <- melt(dcor, 
                  id.vars = c("Grade", "Gender","Age"))

ggplot(df_melted_raw, aes(x = value)) +
  geom_histogram(bins = 100, fill = "orange", color = "black") +
  facet_wrap(~variable, scales = "free")
```

```{r}
#| label: fig-norm-tot
totalNorm = paste0("norm_",unique(df_melted_raw$variable))
totalRaw = as.character(unique(df_melted_raw$variable))
df[,totalNorm] = NA
for(i in 1:length(unique(df_melted_raw$variable))) df[,totalNorm[i]] = normalize(abs(df[,totalRaw[i]]))

df_melted_norm <- melt(df, 
                  id.vars = c("Grade", "Gender", "Code", "Age"),
                  measure.vars = totalNorm)

ggplot(df_melted_norm, aes(x = value)) +
  geom_histogram(binwidth = .5, fill = "orange", color = "black") +
  facet_wrap(~variable, ncol = 3)
```

The correlations are almost identical, however, the time estimation score is now well correlated with the other factors (@fig-norm-cor).

```{r}
#| label: fig-norm-cor
dcor2 <- df[,c(totalNorm,"Grade","Gender")]
corrplot::corrplot(cor(dcor2, use = "pairwise.complete"), addCoef.col = "black")
```

## Factor structure

We may view the different factors measured as indicator of a general factor of time. This might be modeled as a CFA.

```{r}
#| label: general-factor-model
g_model = paste("g =~",paste(totalNorm,collapse="+"))
g_model
fit = cfa(model=g_model, data=df, missing = "fiml",std.lv=T)
fit@Data
fitMeasures(fit,fit.measures=fi)
modificationIndices(fit,sort.=T)[1:10,]
```

The model shows acceptable, yet quite low, fit indices. MI suggest that time discrimination and time reproduction scores may be more similar between each other than with other factors. This makes sense as the two have quite similar requests. With this specification, the model shows excellent fit to the data.

```{r}
#| label: general-factor-model-mi
mi_line <- "
norm_RatioTD	~~	norm_AverageDevAbs_TR
"

g_model <- paste(c(g_model, mi_line), collapse = "\n")
fit = cfa(model=g_model, data=df, missing = "fiml",std.lv=T)
fitMeasures(fit,fit.measures=fi)
```

```{r}
#| label: general-factor-model-mi-summary

summary(fit,std=T)
```

# Time trends

Across primary schools, we expect the time scores to increase, possibly linearly, but probably not. @fig-trend-raw shows the grade trend of the six scales, while @fig-age-trend-raw shows the continuous age trend:

```{r}
#| label: fig-trend-raw
#| warning: false
ggplot(df_melted_raw, aes(y = value, x = Grade)) +
  geom_smooth() + 
  facet_wrap(~variable, ncol = 3, scales = "free")
```

```{r}
#| label: fig-age-trend-raw
#| warning: false
ggplot(df_melted_raw, aes(y = value, x = Age)) +
  geom_smooth() + 
  facet_wrap(~variable, ncol = 3, scales = "free")
```

